---
title: "EXPLORATORY ANALYSIS"
author: "SATYAM SHIVAM SUNDARAM"
output: pdf_document
---

Importing libraries
```{r}
library(tidyverse)
```


Reading data

```{r}
accData <- read_csv('../../data/Accident_Information.csv.bz2')
vehData <- read_csv('../../data/Vehicle_Information.csv.bz2')

```


ACCIDENT DATA
Converting the attributes, those can't contribute to mathematical computations, to factors

```{r}
#names(accData)
accData$Accident_Severity = factor(accData$Accident_Severity,order = TRUE,levels = c("Slight","Serious","Fatal"))
accData[["1st_Road_Class"]] = factor(accData[["1st_Road_Class"]])
accData$Light_Conditions = factor(accData$Light_Conditions)
accData$Day_of_Week = factor(accData$Day_of_Week)
accData$Urban_or_Rural_Area = factor(accData$Urban_or_Rural_Area)
accData$Weather_Conditions = factor(accData$Weather_Conditions)
accData$Junction_Control = factor(accData$Junction_Control)
accData$Road_Surface_Conditions = factor(accData$Road_Surface_Conditions)
accData$Road_Type = factor(accData$Road_Type)
accData$InScotland = factor(accData$InScotland)
accData$Carriageway_Hazards = factor(accData$Carriageway_Hazards)
accData[["2nd_Road_Class"]] = factor(accData[["2nd_Road_Class"]])
accData$Special_Conditions_at_Site = factor(accData$Special_Conditions_at_Site)
accData$Police_Force = factor(accData$Police_Force)
#str(accData)
```


VEHICLE DATA
Converting the attributes, those can't contribute to mathematical computations, to factors

```{r}
#names(vehData)
vehData$Age_Band_of_Driver = factor(vehData$Age_Band_of_Driver,order = "TRUE", levels = c( "Data missing or out of range",
                                                                                           "0 - 5", "6 - 10", "11 - 15", "16 - 20",
                                                                                           "21 - 25", "26 - 35", "36 - 45", 
                                                                                           "46 - 55","56 - 65", "66 - 75", "Over 75"))
vehData$Driver_Home_Area_Type = factor(vehData$Driver_Home_Area_Type)
vehData$Propulsion_Code = factor(vehData$Propulsion_Code)
vehData$Sex_of_Driver = factor(vehData$Sex_of_Driver)
vehData$Skidding_and_Overturning = factor(vehData$Skidding_and_Overturning)
vehData$Vehicle_Manoeuvre = factor(vehData$Vehicle_Manoeuvre)
vehData$X1st_Point_of_Impact = factor(vehData$X1st_Point_of_Impact)
vehData$make = factor(vehData$make)
vehData$Vehicle_Leaving_Carriageway = factor(vehData$Vehicle_Leaving_Carriageway)
vehData$Towing_and_Articulation = factor(vehData$Towing_and_Articulation)
#str(vehData)

```



```{r}

head(accData)
head(vehData)

```

PLOTTING SEVERITY

```{r}

severityData <- subset(accData,select = c("Accident_Index","Accident_Severity"))

cnt <- table(severityData$Accident_Severity)
nd <- as_data_frame(cnt)
names(nd) <- c("Severity","count")
nd

nd$Severity <- factor(nd$Severity)

plt <- ggplot(data = nd,aes(x = Severity,y = count,fill = Severity)) +
        geom_bar(stat="identity") +
        labs(title = "Severity vs count",xlab="Severity",ylab="count")

plt

```





SEVERITY ON THE BASIS OF GENDER OF THE DRIVER

```{r}
severityData2 <- subset(accData,select = c("Accident_Index","Accident_Severity"))

gender2 <- subset(vehData,select = c("Accident_Index","Sex_of_Driver"))%>%
  filter(Sex_of_Driver %in% c("Male","Female"))
severityGender2 <- inner_join(severityData2,gender2)

names(severityGender2) <- c("Index","Severity","Gender")
severityGender2$Gender <- factor(severityGender2$Gender)

plt <- ggplot(severityGender2,aes(x = Severity,fill = Gender))+geom_bar()
plt <- plt + labs(title = "Accident comparison amongst the genders")
plt
  

```







INSIGHTS ABOUT SEVERITY BASED ON AGE-GROUP OF DRIVER AND GENDER

```{r}
severityData3 <- subset(accData,select = c("Accident_Index","Accident_Severity"))

genderAge <- subset(vehData,select = c("Accident_Index","Sex_of_Driver","Age_Band_of_Driver"))%>%
  filter(Sex_of_Driver %in% c("Male","Female") & Age_Band_of_Driver != "Data missing or out of range")

severityGender3 <- inner_join(severityData3,genderAge)

names(severityGender3) <- c("Index","Severity","Gender","Age")
severityGender3$Gender <- factor(severityGender3$Gender)

plt <- ggplot(severityGender3,aes(x = Age,fill = Severity))+geom_bar()+facet_wrap(~Gender)
plt <- plt + labs(title = "Accident comparison amongst the genders of different age groups")
plt


```





```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

